---
title: "lab09:Halloween mini project"
author: "Yufei Zhang A16987415"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

Q1. How many different candy types are in this dataset?

```{r}
nrow(candy)
```
There are 85 types of candy.

Q2. How many fruity candy types are in the dataset?

```{r}
sum(candy[,2])
```
There are 38 fruity candy types.


```{r}
candy["Twix", ]$winpercent
```
Q3. What is your favorite candy in the dataset and what is it’s winpercent value?
```{r}
candy["Sour Patch Kids", ]$winpercent
```

Q4. What is the winpercent value for “Kit Kat”?
```{r}
candy["Kit Kat", ]$winpercent
```

Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?
```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```
```{r}
library("skimr")
skim(candy)
```
Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

All percentages are continuous between 0 and 1 except winpercent looks to be on a different scale. It seems to be in % but not in decimal. All types column are either 0 or 1.

Q7. What do you think a zero and one represent for the candy$chocolate column?

A zero means this candy type does not contain chocolate and a one means it contains chocolate.

Q8. Plot a histogram of winpercent values
```{r}
hist(candy$winpercent)
```

Q9. Is the distribution of winpercent values symmetrical?

No, it is not symmetrical.

Q10. Is the center of the distribution above or below 50%?

It is below 50%

Q11. On average is chocolate candy higher or lower ranked than fruit candy?
```{r}
chocolate<-candy$winpercent[as.logical(candy$chocolate)]
fruity<-candy$winpercent[as.logical(candy$fruity)]
t.test(chocolate,fruity)
```
chocolate candy is ranked higher.

Q12. Is this difference statistically significant?

Yes. Because the t value is 2.87 e-08 which is very very small.

Q13. What are the five least liked candy types in this set?
```{r}
library(dplyr)
head(candy[order(candy$winpercent),], n=5)
candy %>% arrange(winpercent) %>% head(5)
```

Q14. What are the top 5 all time favorite candy types out of this set?
```{r}
candy %>% arrange(desc(winpercent)) %>% head(5)
```
Q15. Make a first barplot of candy ranking based on winpercent values.
```{r}
library(ggplot2)

ggplot(candy) + 
  aes(winpercent, rownames(candy)) +
  geom_col()
```

Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?
```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +  
  geom_col()
```
```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 
```
Now, for the first time, using this plot we can answer questions like:
- Q17. What is the worst ranked chocolate candy?

Sixlets

- Q18. What is the best ranked fruity candy?
Starburst


```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```
Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reese's Miniatures.

Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?
```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```
Nik L Nip is lease popular among the 5.

```{r}
# Make a lollipop chart of pricepercent
ggplot(candy) +
  aes(pricepercent, reorder(rownames(candy), pricepercent)) +
  geom_segment(aes(yend = reorder(rownames(candy), pricepercent), 
                   xend = 0), col="gray40") +
    geom_point()
```

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?
Fruity and chocolate,fruity and bar, pluribus and bar.

Q23. Similarly, what two variables are most positively correlated?

chocolate and winpercent, chocolate and bar.

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```
```{r}
plot(pca$x[, 1:2])
```
```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```
```{r}
# Make a new data-frame with our PCA results and candy data
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```
```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```
```{r}
library(plotly)
ggplotly(p)
```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?
HINT. pluribus means the candy comes in a bag or box of multiple candies.

Fruity and pluribus. Yes since fruity candys usually comes in a bag or box of multiple candies.

